---
title: "PRS Phewas result"
author: X Shen  
date: "\n`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

```{r library,echo=F}
library(dplyr)
library(data.table)
library(readr)
library(pbapply)
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(ggpubr)

source(here::here('scripts/phewas/util/PheWAS_style_p_plot.R'))
```

```{r rendering docs, eval=F,echo=F}
render('scripts/phewas/SUMMARY/summary.prs_phewas.Rmd',output_dir='docs/',output_file='summary.prs_phewas.md')
```



------------------------------------------------------------------------


```{r inputs,echo=F}
f.dat_dir = '/exports/igmm/eddie/GenScotDepression/shen/ActiveProject/Collab/mdd-meta/results/phewas/phewas_out_Body_MRI.rds'
f.dic_dir = '/exports/igmm/eddie/GenScotDepression/shen/ActiveProject/Collab/mdd-meta/results/phewas/data_dictionary/fields.final.brain_imaging_QC_cov_phenotype.txt'
f.category = '/exports/igmm/eddie/GenScotDepression/shen/ActiveProject/Collab/mdd-meta/data/phewas_categories.tsv'
```

```{r load results,echo=F,warning=F}
d.path = f.dic_dir %>% strsplit(.,'/') %>% unlist %>% 
  .[1:(length(.)-1)] %>% 
  paste0(.,collapse = '/')

fields.all = list.files(d.path,pattern='^fields.final',full.names = T,include.dirs = T) %>% 
  as.list %>%
  lapply(fread,stringsAsFactors=F,header=T,sep='\t',quote="") %>%
  lapply(as.data.frame) %>% 
  lapply(mutate,field_used=as.character(field_used),FieldID=as.character(FieldID)) %>% 
  Reduce(function(dtf1,dtf2) bind_rows(dtf1,dtf2[,colnames(dtf1)]), .)


g.path = f.dat_dir %>% strsplit(.,'/') %>% unlist %>% 
  .[1:(length(.)-1)] %>% 
  paste0(.,collapse = '/')

res.targetpT = list.files(g.path,pattern='^phewas_out',full.names = T,include.dirs = T) %>% 
  lapply(readRDS) %>%
  Reduce(function(dtf1,dtf2) bind_rows(dtf1,dtf2), .) %>% 
  mutate(p.corrected = p.adjust(p.value,method='fdr')) %>% 
  mutate(mod_name=as.character(mod_name))

# Find the pT most predictive of MDD CIDI --
tmp.findpt = res.targetpT %>% filter(dependent=='MDD_CIDI') %>% 
  .[grepl('MDD3',.$factor),] %>% 
  .[order(.$beta,decreasing = T),]
pt=tmp.findpt$factor %>% head(n=1) %>% gsub('MDD3_Pt_','',.)
target.pT = pt %>% 
  gsub('\\.','\\\\.',.) %>% 
  paste0('Pt_',.)

res.targetpT = res.targetpT %>% 
  .[grepl(target.pT,.$factor),] 
# --

ref.category = fread(f.category,header=F,sep='\n') %>% 
  as.data.frame %>% .$V1

res.targetpT.annot = res.targetpT %>% 
  left_join(.,fields.all,by=c('dependent'='field_tag')) %>% 
  dplyr::arrange(match(.$category,ref.category)) %>% 
  select(category,Field,field=dependent,factor,beta,std,t.value,p.value,p.corrected)
```

------------------------------------------------------------------------

### Summary

At pT=`r pt`

-   Significant associations with the MDD3 PRS: `r sum(res.targetpT.annot[grepl('MDD3',res.targetpT.annot$factor),'p.corrected']<0.05)`

-   Significant associations with the Howard PRS: `r sum(res.targetpT.annot[grepl('Howard',res.targetpT.annot$factor),'p.corrected']<0.05)`

```{r number of sig assoc, echo=F}
MDD3.nsig = res.targetpT.annot %>%
  filter(grepl('MDD3',factor)) %>% 
  group_by(category) %>%
  summarise(MDD3.Nsig = sum(p.corrected<0.05), Npheno = sum(!is.na(p.corrected)))
Howard.nsig = res.targetpT.annot %>%
  filter(grepl('Howard',factor)) %>%
  group_by(category) %>%
  summarise(Howard.Nsig = sum(p.corrected<0.05))

nsig.compare = MDD3.nsig %>% 
  arrange(match(.$category,ref.category)) %>% 
  left_join(.,Howard.nsig,by='category') %>% 
  select(category,Howard.Nsig,MDD3.Nsig,Npheno)

# nsig.compare %>% 
#   kbl() %>%
#   kable_material(c("striped", "hover"))

```

```{r,fig.height=5,fig.width=10,warning=F,echo=F}
dat.fig.nsig =nsig.compare %>% 
  rename(Category=category) %>% 
  mutate(ord=1:nrow(.))

fig.howard=
  ggplot(dat.fig.nsig, aes(x=reorder(Category,-ord), y=Howard.Nsig)) + 
  geom_bar(fill='darkgoldenrod1',  position=position_dodge(), stat="identity", width=0.8) +
  ggtitle("Howard et al. PRS")+
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line.x = element_line(size=0.5),
    axis.text=element_text(size=8), axis.title=element_text(size=8),
    plot.title = element_text(lineheight=1, face="bold", vjust=1, hjust=0.5,size=9),
    strip.text = element_text(size=8),
    plot.margin=unit(c(1,1,1,3),'mm')) +
  ylab("No. of significant associations") + xlab("\n\n") +
  #scale_y_continuous(limits=c(0,57))+
  scale_y_reverse(lim=c(125,0))+
  scale_x_discrete(position='top')+
  geom_hline(yintercept=0,size=0.5,color='grey')+
  coord_flip()

fig.mdd3=
  ggplot(dat.fig.nsig, aes(x=reorder(Category,-ord), y=MDD3.Nsig)) + 
  geom_bar(fill='indianred1',  position=position_dodge(), stat="identity", width=0.8) +
  ggtitle("MDD3 PRS")+
   theme(
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_text(hjust=0.5),
        axis.line.x = element_line(size=0.5),
        axis.text=element_text(size=8), axis.title=element_text(size=8),
        plot.title = element_text(lineheight=1, face="bold", vjust=1, hjust=0.5,size=9),
        strip.text = element_text(size=8)) +
  ylab("No. of significant associations")  +
  geom_hline(yintercept=0,size=0.5,color='grey')+
  ylim(0,125) +
  coord_flip()

Fig.total=ggarrange(fig.howard,fig.mdd3,
                 widths = c(1,1.8),
                 nrow = 1, ncol=2, align='h')

ggsave(Fig.total,filename = here::here('results/phewas/figs/Nsig_Howard_MDD3.png'),width = 10,height = 4,units = 'in',device = 'png')

Fig.total

```


------------------------------------------------------------------------

### P-plots

#### MDD3

```{r}
ref.two_level.category = data.frame(lv2=ref.category,stringsAsFactors = F) %>% 
  mutate(lv1=ifelse(grepl('Mental health',lv2),'Mental health',
                    ifelse(grepl('Early life|Socio|Lifestyle|Diet|Physical activity',lv2),'Environment',
                           ifelse(grepl('Physical health',lv2),'Self-reported physical condition',
                    ifelse(grepl('Physical|Blood|Body MRI',lv2),'Physical assessment',
                           'Brain and cognition')))))

```


```{r pplot mdd3,echo=F,warning=F,fig.height=16,fig.width=18}
res.targetpT.annot.forplot = res.targetpT.annot %>% 
  mutate(category=gsub('Desikan atlas ','',category)) %>% 
  mutate(category=gsub('atlas ','',category)) %>% 
  mutate(category=gsub('tissue measure','tissue',category)) %>% 
  mutate(Field=gsub(' (left hemisphere)','',Field,fixed = T)) %>% 
  mutate(Field=gsub(' (whole brain)','',Field,fixed = T)) %>% 
  mutate(Field=gsub('_',' ',Field,fixed = T)) %>% 
  mutate(Field=gsub('medialorbitofrontal','medial orbito-frontal',Field,fixed = T))
  
# wrap text if too long: category and Field
res.targetpT.annot.forplot$category = res.targetpT.annot.forplot$category %>% 
  as.list %>% lapply(.,strwrap,width=25) %>% 
  lapply(.,paste0,collapse='\n') %>% 
  unlist

res.targetpT.annot.forplot$Field = res.targetpT.annot.forplot$Field %>% 
  as.list %>% lapply(.,strwrap,width=25) %>% 
  lapply(.,paste0,collapse='\n') %>% 
  unlist
  
plot.dat = res.targetpT.annot.forplot %>%
  filter(grepl('MDD3',factor)) %>% 
  rename(dependent=field) %>% 
  left_join(.,ref.two_level.category,by=c('category'='lv2'))


ls.label = plot.dat %>%
  filter(p.corrected<0.05) %>%
  arrange(p.value) %>% 
  group_by(lv1) %>% slice(1:3) %>% 
  arrange(match(category,ref.category)) %>% 
  as.data.frame %>% 
  select(dependent,Field) %>% 
  .[!duplicated(.$Field),]

source(here::here('scripts/phewas/util/PheWAS_style_p_plot_poster.R'))

fig.MDD3=p_plot(TargetResult=plot.dat,color_theme_usr='Shen',
                shape_sig=T,sig_nominal=F,
           plot_title=paste0('MDD3 PRS (pT=',pt,')'),y_lim=310,
           labels_annot=ls.label,add_category_name=T)

#fig.MDD3

ggsave(fig.MDD3,filename = here::here('results/phewas/figs/pplot_MDD3.png'),width = 17,height = 10,units = 'in',device = 'png')
```

#### Howard et al

```{r pplot Howard,echo=F,warning=F,fig.height=16,fig.width=18}
plot.dat = res.targetpT.annot.forplot %>%
  filter(grepl('Howard',factor)) %>% 
  rename(dependent=field)


ls.label = plot.dat %>%
  filter(p.corrected<0.05) %>%
  arrange(p.value) %>% 
  group_by(category) %>% slice(1:5) %>% 
  arrange(match(category,ref.category)) %>% 
  as.data.frame %>% 
  select(dependent,Field)
#source(here::here('scripts/phewas/util/PheWAS_style_p_plot.R'))

fig.Howard=p_plot(TargetResult=plot.dat,color_theme_usr='Shen',
                shape_sig=T,sig_nominal=F,
           plot_title=paste0('Howard PRS (pT=',pt,')'),y_lim=310,
           labels_annot=ls.label,add_category_name=T)

fig.Howard
ggsave(fig.Howard,filename = here::here('results/phewas/figs/pplot_Howard.png'),width = 30,height = 15,units = 'in',device = 'png')
```

