---
title: Conditional and Joint Analysis
output: github_document
---

```{r warning=FALSE, message=FALSE}
library(rtracklayer)
library(readxl)
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(UpSetR)
library(genpwr)
library(ggplot2)
```


```{r echo=FALSE}
qc <- snakemake@params$qc
```

# Methods

We ran a [conditional and joint analysis](https://www.nature.com/articles/ng.2213) using [GCTA](https://cnsgenomics.com/software/gcta/#COJO) to refine the list of independent loci.

- Final meta-analysed SNPs from the Ricopili pipeline were used.
- Ricopili was used for initial clumping with index SNPs identified with $p <$ `r qc$clu_p1` and $r^2 <$ `r qc$clu_r2` within `r qc$clu_kb`kb windows. The extended MHC region was clumped as a single region.
- Sumstats were filtered for MAF >= `r qc$clu_maf`, INFO > `r qc$clu_info`, and Neff >= 80% of max.
- Regions with a genome-wide significant SNP (p < 5-e8) were identified from the clumped results. Regions within `r qc$cojo_kb`kb of each other were merged.
- SNPs from these regions were extracted and filtered to unrelated of self- and genotype-identified European ancestry participants from UK Biobank.
- A conditional analysis was performed on each region using the filtered sumstats superimposed on the UK Biobank LD structure.
- Singleton regions (with only one genome-wide significant variant) were removed.

## Previous sumstats

Variants from [Wray et al 2018](https://www.nature.com/articles/s41588-018-0090-3/tables/2), [Howard et al 2019](https://www.nature.com/articles/s41593-018-%200326-7) (remove first and last rows with table captions), and [Levey et al 2021](https://doi.org/10.1038/s41593-021-00860-2). Parse out regions from Wray and Howard.

```{r}
wray <- read_tsv(snakemake@input$wray) %>%
    separate(`Region (Mb)`, into=c('range.left.Mb', 'range.right.Mb'), sep='–', convert=TRUE) %>%
    mutate(range.left=range.left.Mb*1e6, range.right=range.right.Mb*1e6)

howard <- read_excel(snakemake@input$howard, skip=2, n_max=102) %>%
    separate(`Region (bp) of clumped variants (P < 10-4) in linkage disequilibrium (r2 > 0.1) with lead variant`, into=c('range.left', 'range.right'), sep='-', convert=TRUE)

levey <- read_tsv(snakemake@input$levey, col_types=cols(CHR.BP=col_character()))
```

# Results

## COJO SNP and region counts

List of clumped and COJO SNPs and regions

```{r echo=F, results='asis'}
cat(paste('- ', readLines(snakemake@input$log)[-1], '\n'))
```

Load list of COJO SNPs
```{r}
cojo <- read_tsv(snakemake@input$cojo)
```

## Clumped results

Clumped results from Ricopili

```{r}
rp <- read_table2(snakemake@input$rp_clump) %>% filter(P <= 5e-8)
```

## Genomic ranges

Construct genomic range objects so that SNPs and regions can be intersected and compared.

```{r}
cojo_gr <- with(cojo, GRanges(seqnames=CHR, ranges=IRanges(start=range.left, end=range.right), SNP=SNP))
rp_gr <- with(rp, GRanges(seqnames=CHR, ranges=IRanges(start=range.left, end=range.right), SNP=SNP))

wray_gr <- with(wray, GRanges(seqnames=Chr., ranges=IRanges(start=range.left, end=range.right, SNP=SNP)))
howard_gr <- with(howard, GRanges(seqnames=Chromosome, ranges=IRanges(start=range.left, end=range.right, SNP=`Marker Name`)))
levey_gr <- with(levey, GRanges(seqnames=CHR, ranges=IRanges(start=BP, width=1)))
```

## GWAS catalog

Tally entries in the GWAS catalog for each region

```{r}

parse_catalog_entry <- function(catalog_entry) {
    # extract LD and RSID from first element
    ld_snp_match <- str_match(catalog_entry[1], "\\((.+)\\)(rs[[:digit:]]+)")
    ld <- as.numeric(ld_snp_match[,2])
    snp <- ld_snp_match[,3]
    # parse rest of elements into phenotype(PubMed ID)(P-value)
    phenotype_matches <- str_match(catalog_entry[-1], "(.+)\\(([[:digit:]]+)\\)\\((.+)\\)")
    phenotypes <- phenotype_matches[,2]
    pubmeds <- as.numeric(phenotype_matches[,3])
    p_values <- as.numeric(phenotype_matches[,4])
    
    return(data.frame(ld, catalogSNP=snp, phenotype=phenotypes, pubmed_id=pubmeds, P=p_values))
}


rp_gwas_catalog_entries <-
plyr::ddply(filter(rp, gwas_catalog_span.6 != '-'), ~SNP, function(rp_entry) {
    # get GWAS catalog cell
    gwas_catalog <- rp_entry$gwas_catalog_span.6
    # split out into catalog entries (separated by /, removing first empty element)
    catalog_entries <- str_split(gwas_catalog, "/")[[1]]
    catalog_entries_complete <- catalog_entries[which(catalog_entries != "")]
    # split SNP entries by ";"
    catalog_entries_list <- str_split(catalog_entries_complete, ";")
    # parse each entry
    catalog_entries_df <- plyr::ldply(catalog_entries_list, parse_catalog_entry)
    return(catalog_entries_df)
}) %>% as_tibble()

```

# Parse gene list

```{r}

rp_genes_dist <- 
plyr::ddply(filter(rp, `genes.6.50kb(dist2index)` != '-'), ~SNP, function(rp_entry) {
    genes_dist <- rp_entry$`genes.6.50kb(dist2index)`
    genes_dist_list <- str_split(genes_dist, ',')[[1]]
    genes_dist_match <- str_match(genes_dist_list, "(.+)\\((.+)\\)")
    return(data.frame(gene=genes_dist_match[,2], dist2index=as.numeric(genes_dist_match[,3])))
}) %>% as_tibble()

```

## Comparison to previous findings

Find overlaps between clumped, COJO, and previous results. Append and then reduce all regions

```{r}

all_gr <- reduce(c(cojo_gr, rp_gr, wray_gr, howard_gr, levey_gr))

```

Find overlaps and make lists for an upset plot. Take the index from the combined set as the element, then find which of them are found within each of the sets of SNPs. The upset plot function handles finding each combination of intersections.

```{r upset, fig.width=10, fig.height=8}

hits_upset <- list(COJO=unique(findOverlaps(all_gr, cojo_gr)@from),
                   Clump=unique(findOverlaps(all_gr, rp_gr)@from),
                   Wray=unique(findOverlaps(all_gr, wray_gr)@from),
                   Howard=unique(findOverlaps(all_gr, howard_gr)@from),
                   Levey=unique(findOverlaps(all_gr, levey_gr)@from))
                   
upset(fromList(hits_upset), order.by='freq', text.scale=2)

```

Find which COJO regions overlap with Howard

```{r}
cojo_howard_overlaps <- findOverlaps(cojo_gr, howard_gr)
cojo_howard_overlaps
```

Count number of regions in Howard that overlap:

```{r}
howard %>% slice(unique(cojo_howard_overlaps@to)) %>% count()
```

Find which COJO regions overlap with Levey

```{r}
cojo_levey_overlaps <- findOverlaps(cojo_gr, levey_gr)
cojo_levey_overlaps
```

Count number of regions in Levey that overlap:

```{r}
levey %>% slice(unique(cojo_levey_overlaps@to)) %>% count()
```

```{r}
cojo_known <- cojo %>% slice(unique(cojo_levey_overlaps@from))

catalog_known <- rp_gwas_catalog_entries %>% filter(SNP %in% cojo_known$SNP) %>% count(phenotype) %>% arrange(desc(n))
catalog_known
```

Newly discovered regions

```{r}
cojo_new <- cojo %>% slice(unique(-cojo_levey_overlaps@from)) %>% arrange(P)
cojo_new %>%
select(region, snp_idx, CHR, SNP, BP, P, pJ) %>%
group_by(region)
```

```{r}
rp_gwas_catalog_entries %>% filter(SNP %in% cojo_new$SNP) %>% count(phenotype) %>% arrange(desc(n)) %>% filter(!phenotype %in% catalog_known$phenotype)
```

```{r}
rp_genes_dist %>% filter(SNP %in% cojo_new$SNP) %>% group_by(SNP) %>% filter(dist2index == min(dist2index)) %>% ungroup() %>% select(gene) %>% distinct()
```

Calculate power for previous versus current GWAS

```{r}

# use sum of effective sample sizes and case rate of 50% rather than sum of cases and controls
levey_power <- genpwr.calc(calc='es', model='logistic', ge.interaction=NULL, N=938012, Case.Rate=0.5, k=NULL, MAF=seq(0.005, 0.49, by=0.005), Power=0.8, Alpha=5e-8, True.Model='Additive', Test.Model='Additive')

mdd2022_power <- genpwr.calc(calc='es', model='logistic', ge.interaction=NULL, N=2*776935, Case.Rate=0.5, k=NULL, MAF=seq(0.005, 0.49, by=0.005), Power=0.8, Alpha=5e-8, True.Model='Additive', Test.Model='Additive')

```

```{r cojo_known_novel}

frq_u_col <- str_subset(names(cojo), 'FRQ_U')

cojo_known_novel <- bind_rows(
mutate(cojo_known, assoc='Known'),
mutate(cojo_new, assoc='Novel')) %>%
mutate(BETA=log(OR)) %>%
select(SNP, assoc, OR, BETA, FRQ=starts_with('FRQ_U')) %>%
mutate(MAF=if_else(FRQ <= 0.5, true=FRQ, false=1-FRQ))

power_known_novel <- bind_rows(
transmute(levey_power, Power='Levey', MAF, OR=`OR_at_Alpha_5e-08`),
transmute(mdd2022_power, Power='MDD2022', MAF, OR=`OR_at_Alpha_5e-08`)
)

ggplot(cojo_known_novel, aes(x=MAF, y=exp(abs(BETA)))) + 
geom_point(aes(color=assoc)) +
geom_line(mapping=aes(y=OR, linetype=Power), data=power_known_novel, size=1) +
scale_y_continuous('OR', limits=c(1, 1.1))

```

## Comparison between pre-COJO and post-COJO

Find regions overlapping between clumped and COJO results
```{r}

cojo_clumped_overlaps <- findOverlaps(cojo_gr, rp_gr)

```

List COJO results where the selected SNP is not in the clumped results
```{r}
cojo_newly_selected <- 
cojo %>% slice(unique(cojo_clumped_overlaps@from)) %>%
filter(!SNP %in% rp$SNP) %>%
select(region, snp_idx, CHR, SNP, P, pJ) 
cojo_newly_selected
```

Newly selected SNPs that were not GWsig in the clumped results
```{r}
cojo_newly_selected %>%
filter(P > 5e-8)
```

List clumped SNPs in retained regions that were not selected by COJO
```{r}
rp %>% slice(unique(cojo_clumped_overlaps@to)) %>%
filter(!SNP %in% cojo$SNP)
```


Line up non-selected SNPs with selected SNPs in the region
```{r}
bind_cols(
select(slice(cojo, cojo_clumped_overlaps@from), region, snp_idx, SNP.cojo=SNP, BP.cojo=BP, P.cojo=P, PJ.cojo=pJ),
select(slice(rp, cojo_clumped_overlaps@to), SNP.rp=SNP, BP.rp=BP, P.rp=P)
) %>%
filter(!SNP.rp %in% cojo$SNP)
```



